Comparative vaccine effectiveness of ChAdOx1 versus BNT162b2 in Health and Social Care workers in England

### 2021-09-10 analysis/report/draft-manuscript.Rmd
add second dose to irr table; remove fu time from table1;

data_cohort <- 
  data_cohort %>%
  mutate(
    censor_date = pmin(vax1_date - 1 + lastfupday, end_date, dereg_date, death_date, vax2_date, na.rm=TRUE),
    tte_censor = tte(vax1_date-1, censor_date, censor_date, na.censor=TRUE),

    postest = censor_indicator(positive_test_date, censor_date),
    emergency = censor_indicator(emergency_date, censor_date),
    covidemergency = censor_indicator(emergency_covid_date, censor_date),
    admitted = censor_indicator(admitted_date, censor_date),
    covidadmitted = censor_indicator(covidadmitted_date, censor_date),
    covidcc = censor_indicator(covidcc_date, censor_date),
    coviddeath = censor_indicator(coviddeath_date, censor_date),
    death = censor_indicator(death_date, censor_date),
  )

baseline <- 
  data_cohort %>%
  summarise(
    n = n(),
    age_median = median(age),
    age_Q1 = quantile(age, 0.25),
    age_Q3 = quantile(age, 0.75),
    female = mean(sex=="Female"),
    fu_years = sum(tte_censor)/365.25,
    fu_median = median(tte_censor),
    priorinfection = mean(prior_covid_infection),

    postest = sum(postest),
    covidemergency = sum(covidemergency),
    emergency = sum(emergency),
    admitted = sum(admitted),
    covidadmitted = sum(covidadmitted),
    coviddeath = sum(coviddeath),
    death = sum(death),
  )

baseline_az <- 
  data_cohort %>%
  filter(vax1_type=="az") %>%
  summarise(
    n = n(),
    age_median = median(age),
    age_Q1 = quantile(age, 0.25),
    age_Q3 = quantile(age, 0.75),
    female = mean(sex=="Female"),
    fu_years = sum(tte_censor)/365.25,
    fu_median = median(tte_censor),
    priorinfection = mean(prior_covid_infection),
    median_vaxday = first(start_date_az) + floor(median(vax1_day)) -1
  )

baseline_pfizer <- 
  data_cohort %>%
  filter(vax1_type=="pfizer") %>%
  summarise(
    n = n(),
    age_median = median(age),
    age_Q1 = quantile(age, 0.25),
    age_Q3 = quantile(age, 0.75),
    female = mean(sex=="Female"),
    fu_years = sum(tte_censor)/365.25,
    fu_median = median(tte_censor),
    priorinfection = mean(prior_covid_infection),
    median_vaxday = first(start_date_az) + floor(median(vax1_day)) - 1
  )
```

@@

@@

This observational study of almost 1/3 million health and social care workers living in England found no significant differences in the incidence of COVID-19-related events between those receiving BNT162b2 or ChAdOx1 at 6 weeks. There is a clear leveling-off of the incidence of positive tests after around 3 weeks, consistent with the expected time-to-onset of vaccine-induced immunity of around 2 weeks plus the delay from infection to a positive test. After

@@

>>>>

data_cohort <- 
  data_cohort %>%
  mutate(
    censor_date = pmin(vax1_date - 1 + lastfupday, end_date, dereg_date, death_date, na.rm=TRUE),
    tte_censor = tte(vax1_date-1, censor_date, censor_date, na.censor=TRUE),
    tte_seconddose = tte(vax1_date-1, vax2_date-1, censor_date, na.censor=TRUE),

    postest = censor_indicator(positive_test_date, censor_date),
    emergency = censor_indicator(emergency_date, censor_date),
    covidemergency = censor_indicator(emergency_covid_date, censor_date),
    admitted = censor_indicator(admitted_date, censor_date),
    covidadmitted = censor_indicator(covidadmitted_date, censor_date),
    covidcc = censor_indicator(covidcc_date, censor_date),
    coviddeath = censor_indicator(coviddeath_date, censor_date),
    death = censor_indicator(death_date, censor_date),

    seconddose = censor_indicator(vax2_date-1, censor_date)
  )

baseline <- 
  data_cohort %>%
  summarise(
    n = n(),
    age_median = median(age),
    age_Q1 = quantile(age, 0.25),
    age_Q3 = quantile(age, 0.75),
    female = mean(sex=="Female"),
    fu_years = sum(tte_censor)/365.25,
    fu_median = median(tte_censor),
    priorinfection = mean(prior_covid_infection),
    seconddose_12week = mean(replace_na(tte_seconddose<=12*7, FALSE)),

    postest = sum(postest),
    covidemergency = sum(covidemergency),
    emergency = sum(emergency),
    admitted = sum(admitted),
    covidadmitted = sum(covidadmitted),
    coviddeath = sum(coviddeath),
    death = sum(death),
    seconddose = sum(seconddose),
  )

baseline_az <- 
  data_cohort %>%
  filter(vax1_type=="az") %>%
  summarise(
    n = n(),
    age_median = median(age),
    age_Q1 = quantile(age, 0.25),
    age_Q3 = quantile(age, 0.75),
    female = mean(sex=="Female"),
    fu_years = sum(tte_censor)/365.25,
    fu_median = median(tte_censor),
    priorinfection = mean(prior_covid_infection),
    median_vaxday = first(start_date_az) + floor(median(vax1_day)) -1,
    seconddose_12week = mean(replace_na(tte_seconddose<=12*7, FALSE)),

  )

baseline_pfizer <- 
  data_cohort %>%
  filter(vax1_type=="pfizer") %>%
  summarise(
    n = n(),
    age_median = median(age),
    age_Q1 = quantile(age, 0.25),
    age_Q3 = quantile(age, 0.75),
    female = mean(sex=="Female"),
    fu_years = sum(tte_censor)/365.25,
    fu_median = median(tte_censor),
    priorinfection = mean(prior_covid_infection),
    median_vaxday = first(start_date_az) + floor(median(vax1_day)) - 1,
    seconddose_12week = mean(replace_na(tte_seconddose<=12*7, FALSE)),
  )
```

@@

At 12 weeks, `r label_percent(0.1)(baseline_pfizer$seconddose_12week)` of BNT162b2 recipients had received a second dose, compared with `r label_percent(0.1)(baseline_az$seconddose_12week)` of ChAdOx1 recipients.

@@

This observational study of almost 1/3 million health and social care workers living in England found no significant differences in the incidence of COVID-19-related events between those receiving BNT162b2 or ChAdOx1 at 6 weeks. There is a clear leveling-off of the incidence of positive tests after around 3-4 weeks, consistent with the expected time-to-onset of vaccine-induced immunity of around 2 weeks plus the delay from infection to a positive test.

@@

Second-dose effects were considered using an intention-to-treat approach, which does not account for potential differences in treatment adherence between vaccine types. However, we found that by 12 weeks more BNT162b2 recipients had received a second dose than ChAdOx1 recipients, which would bias effect estimates in favour of BNT162b2 assuming higher protection after a second-dose of either vaccine. Either way, there was insufficient power to reliably estimate comparative second-dose effects as event incidence at 12 weeks and beyond was low for both vaccines, with fewer than 10 positive tests per 1000-person-years, and far fewer for COVID-19 A&E attendances and hospital admissions during this period. The need for many thousands of second-dose person-years to detect differences in second-dose effects is encouraging to the extent that it suggests extremely high protective effects for both vaccines, though the fact that background case rates in England were declining over the duration of the study period should not be ignored <!-- covid dashboard -->. We also cannot rule-out potential waning effects, which we did not study.
